Random Fibonacci Anyons Simulable via Covariant Action of Mapping Class Group

### 2014-08-29 ising/Fibonacci/about.tex

\heading{Fibonacci Anyons}

Every surface that is not a sphere, torus, disc or annulus, admits
a pair-of-pants decomposition, see \cite{Ivanov01}, Theorem 2.4.A.

@@

Where $F_n$ is the Hilbert state space of $n$ Fibonacci anyons with
trivial total charge.

@@

We aim discretize the problem for software implementation.
Our first task is to characterize what happens to a curve
if we ``forget'' any braiding inside a disc.
We define the following set of ``planar diagrams'' as:

@@

where $f:L\to D$ is a curve (not an isotopy class)
whose image $f(L)$ intersects a disc $E\subset D$ in a finite number of components
and the equivalence relation $\sim$ consists of isotopies that fix $\partial E.$

This object has a simple combinatorial description 
as a string consisting of
the four labels {\bf ( ) H T} such that 
the parentheses are balanced. The two labels H and T specify the
head and tail endpoints of $f$, and each appears at most once in the string.

@@

See \cite{Abramsky08} Proposition 1.6, for a similar treatment.

@@

Now we consider a tiling of the disc into subdiscs that are disjoint on
their interiors.

>>>>

\heading{Anyons}

We measure charge contained within a disc. % region, circle
Commuting measurements are either disjoint discs or
one wholly contained within the other:

So (proof?) we get a basis for the Hilbert space from
a pair-of-pants (POP) decomposition:

Every surface that is not a sphere, torus, disc or annulus, has
a POP decomposition, see \cite{Ivanov01}, Theorem 2.4.A.

We say that a POP decomposition of $D_n$
{\it admits} a curve $f:L_n\to D_n$
if the curve performs a depth first traversal of $Q_n.$
This means that if the curve visits the trunk of a POP
then it visits all the legs of the POP before leaving the POP.

Conversely, given a curve $f:L_n\to D_n$
and a binary rooted tree $t$,
we can recover a unique POP that admits $f$.
(proof: use small neighbourhoods of the curve)

\important{F-moves:} given a nested POP and admissable $f$,
the f-move is the change of basis as follows:

\important{R-moves:} we examine the action of $MCG(D_n)$
on our state space. Given a single POP and curve $f:L_2\to D_n$,
the half-twist acts on $\psi$ as $R\psi$.

This together with f-moves gives all the half-twists on
the curve $f$.

@@

Where $F_n$ is the Hilbert state space of $n$ Fibonacci anyons.
%with trivial total charge.

See \cite{Pfeifer12, Pfeifer14}

@@

We aim discretize the problem for software implementation.
Our first task is to characterize what happens to a curve
if we ``forget'' any braiding inside a disc.
We adapt the Abramsky combinatorial formalism 
of the Temperly-Lieb algebra to our needs (\cite{Abramsky08}, Proposition 1.6).
The following set of ``planar diagrams'' as is defined as:

@@

where $f:L\to D$ is a curve (not an isotopy class)
whose image $f(L)$ intersects a disc $E\subset D$ in a finite number of components
and the equivalence relation $\sim$ consists of 
isotopies that fix $\partial E.$ % Local (relative?) isotopies

This object has a simple combinatorial description 
as a string consisting of
the four labels {\bf (, ), H, T} such that 
the parentheses are balanced. The two labels {\bf H} and {\bf T} specify the
head and tail endpoints of $f$, each of which appears at most once in the string.

@@

@@

%Now we consider a tiling of the disc into subdiscs that are disjoint on
%their interiors.
We consider now a cellulation of a disc.
Concretely, we choose a regular $l$ by $l$ square tiling of a square region.

The holes $Q_n$ will always be contained within the interior of the
tiles.

\important{Definition:} the discrete curve space $C_n^(l)$ consists of
curves $L_n\to D_n$ modulo isotopy that fixes the boundary of
every tile.
